---
title: "p8105_hw3_ah4167"
author: "Aiying Huang"
date: "2023-10-05"
output: github_document
---

```{r}
library(tidyverse)
```
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```


## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```
```{r}
instacart|>
  group_by(aisle)|>
  summarize(n_products=n())
```
```{r}
instacart|>
  group_by(aisle)|>
  summarize(n_products=n())|>
  mutate(
    products_rank=min_rank(desc(n_products)))|>
  filter(products_rank==1)
```

There are `r nrow(instacart|>group_by(aisle)|>summarize(n_products=n()))` aisles , and `r instacart|>group_by(aisle)|>summarize(n_products=n())|>mutate(products_rank=min_rank(desc(n_products)))|>filter(products_rank==1)|>pull(aisle)` are the most items ordered from.


Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart|>
  group_by(aisle)|>
  summarize(n_products=n())|>
  filter(n_products>10000)|>
  ggplot(aes(x=aisle,y=n_products))+
  geom_point()+
  labs(
    title=" Items ordered in each aisle",
    x="Aisle Name",
    y="Numbers of items"
  )+
  theme_minimal()
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
baking_ingredients =
  instacart |>
  filter(
    aisle == "baking ingredients") |>
  group_by(product_name)|>
  summarize(order_count=n())|>
  mutate(
    products_rank=min_rank(desc(order_count)),
    aisle_type="baking ingredients")|>
  filter(products_rank<4)
dog_food_care =
  instacart |>
  filter(
    aisle == "dog food care") |>
  group_by(product_name)|>
  summarize(order_count=n())|>
  mutate(
    products_rank=min_rank(desc(order_count)),
    aisle_type="dog food care")|>
  filter(products_rank<4)
packaged_vegetables_fruits =
  instacart |>
  filter(
    aisle == "packaged vegetables fruits") |>
  group_by(product_name)|>
  summarize(order_count=n())|>
  mutate(
    products_rank=min_rank(desc(order_count)),
    aisle_type="packaged vegetables fruits"
    )|>
  filter(products_rank<4)
```
```{r}
  rbind(baking_ingredients, dog_food_care, packaged_vegetables_fruits)|>
  arrange(aisle_type,products_rank)|>
  relocate(aisle_type,products_rank,product_name)|>
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
Pink_Lady_Apples_df=
instacart|>
  group_by(product_name)|>
  filter(product_name=="Pink Lady Apples")|>
  select(order_dow,order_hour_of_day,product_name)|>
  group_by(order_dow)|>
  summarize(
    mean_day_hr=mean(order_hour_of_day)
  )|>
  mutate(
    product_name="Pink Lady Apples"
  )
Coffee_Ice_Cream_df=
  instacart|>
  group_by(product_name)|>
  filter(product_name=="Coffee Ice Cream")|>
  select(order_dow,order_hour_of_day,product_name)|>
  group_by(order_dow)|>
  summarize(
    mean_day_hr=mean(order_hour_of_day)
  )|>
  mutate(
    product_name="Coffee Ice Cream"
  )
```
```{r}
rbind(Pink_Lady_Apples_df, Coffee_Ice_Cream_df)|>
  mutate(
    order_dow = case_match(
      order_dow,
      0 ~ "Sunday",
      1 ~ "Monday",
      2 ~ "Tuesday",
      3 ~ "Wednesday",
      4 ~ "Thursday",
      5 ~ "Friday",
      6 ~ "Saturday"
  ))|>
  pivot_wider(
    names_from=order_dow,
    values_from=mean_day_hr
  )|>
  knitr::kable()
```

## Problem 2

```{r}
data("brfss_smart2010")
```

First, do some data cleaning:

```{r}
brfss_clean_df=
brfss_smart2010|>
  janitor::clean_names()|>#format the data to use appropriate variable names
  filter(
    topic=="Overall Health",#focus on the “Overall Health” topic
    response %in% c("Excellent", "Very Good", "Good", "Fair", "Poor")#include only responses from “Excellent” to “Poor”
  )|>
  mutate(
    response=factor(
      response,
      levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"), 
      ordered = TRUE
    )#organize responses as a factor taking levels ordered from “Poor” to “Excellent”
  )|>
  separate(locationdesc, into = c("state", "location"),sep=" - ")|>
  select(year,state,location,class,topic,question,response,sample_size,data_value,geo_location)
```

Then we can calculate information below:

In 2002, the states observed at 7 or more locations are `r brfss_clean_df|>filter(year==2002)|>group_by(state,location)|>summarize(response_count=n())|>summarize(location_count=n())|>filter(location_count>=7)|>pull(state)`. 
In 2010, the states observed at 7 or more locations are `r brfss_clean_df|>filter(year==2010)|>group_by(state,location)|>summarize(response_count=n())|>summarize(location_count=n())|>filter(location_count>=7)|>pull(state)`.

Furthermore, we construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.

```{r}
#Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
brfss_clean_df|>
  filter(response=="Excellent")|>
  group_by(state,year)|>
  summarize(avg_data_value=mean(data_value,na.rm=TRUE))|>
#Make a “spaghetti” plot of this average value over time within a state
  ggplot(aes(x = year, y = avg_data_value, color = state, group = state)) +
  geom_line() +
  labs(title = "Spaghetti Plot of Average Value Over Time Within a State",
       x = "Year",
       y = "Average Value",
       color = "State") +
  theme_minimal()
```

--interpretation to be completed：We can see from the plot that 

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss_clean_df|>
  filter(year%in%c(2006,2010)&state=="NY")|>
  ggplot(aes(x=response,y=data_value))+
  geom_boxplot()+
  facet_wrap(~year)+
  labs(
    title="Distribution of Data Value by Response in NY State",
    x = "Response",
    y = "Data Value"
  )
```

--interpretation to be completed：We can see from the plot that 














